Peptide retention time prediction

### 2014-12-04 retention.tex
tired

% This paragraph is consfusing, should be rephrased
The principle of predictability also extends to targeted proteomics,
in SRM analysis the peptides retention time and fragments are
predicted from retention time, often derived from previous shotgun
analysis. As targeted proteomics move into more comprehensive
techniques, such as DIA/SWATH analysis, it becomes increasingly
attractive to use techniques that do not rely on previous observations
of the exact peptide to determine its properties, but instead to use
de-novo predictions from amino acid sequences.

% New technology to learn 
% - ranking-based method
% - deep learning

>>>>

Moreover, targeted proteomics experiments, such as Selected Reaction
Monitoring (SRM) experiments, are dependent on the predictability of a
peptides behavior. In such experiments the retention times and the
suitable transitions are needed as input for the analysis. Frequently,
such information is derived from previous shotgun analysis. However,
as the capacity of targeted proteomics experiments expands by the
introduction of the more comprehensive data independent acquisition
techniques techniques (a.k.a. SWATH analysis), the pre-identification
of peptide transitions and retention times threatens to become a
limiting factor, as the analysis never can become more comprehensive
than the previous shotgun analysis would be. In such a scenario it
becomes increasingly attractive to predict retention time and
transition using {\em de novo} predictions from the peptides amino
acid sequences.  A system of identifications not dependent on previous
peptide identifications from shotgun proteomics experiment would lead
to a more direct experimental design and hopefully to more
comprehensie coverage. Thus {\em de novo} retention time predictions
have an important role to fill. 
%However, to do so we need to make the technique more accurate.

We can see that there are a number of interesting machine learning
techniques that not yet have been used on the problem of predicting
peptides retention time. Here we will bring up one such techniques,
that we feel are particularly good matches, that we have not yet seen
implemented for the problem. Pairwise Machine-learned ranking (MLR)
(\url{http://dx.doi.org/10.1561%2F1500000016}) models can be trained
to determine which of two examples that should be better ranked thatn
the other. MLR is traditionally used for construction of ranking
models for information retrieval systems. However, in the case of
retention time prediction we can use such a predictor to predict the
elusion order for sets of peptides, which can be useful in its own
(coon; just in time), but can also be converted into absolute
retention times, given that the distribution of their retention times
are known.

It is also worth noting that the current trend of processing huge data
repositories, frequently referred to as bigdata, have put quite some
intest in so called deep learning techniques such as belief networks
using restricted multiple layered bolzman machines (RBM; cite
http://bit.ly/1vMq6az) or deep neural networks. The idea is that the
performance of predictors often are enhanced if one can structure the
prediction into multiple layers. The drawback of such models are that
they require large training seta. Given the very large sets of
peptides with retention times that are assembled in various
repositories, such as peptide atlas and pride, such methods could
probably be implemented for retention time prediction as well.
